---
title: "Diet power analysis"
author: "Rosie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
library(tidyverse)
library(lubridate)
library(lme4)
library(lmerTest)
library(simr)
```

## R Markdown

Let's look at stomach fullness and diet weight variability to see if we can use it in 2023

```{r}
load("Dietdata.Rdata")

dietsmelt = select(diet, Date, Cage.ID, Location, Tag, Fullness, Digestion, Contents.Weight, Season) %>%
  distinct() %>%
  mutate(smeltID = paste(Location, Date, Tag), Month = month(Date), CageID2 = paste(Location, Cage.ID)) 
  

ggplot(dietsmelt, aes(x = Fullness))+ geom_density()+
  facet_wrap(~Location)

```

Some exploritory plots


```{r}
ggplot(dietsmelt, aes(x = smeltID, y = Contents.Weight))+ geom_col()+
  facet_wrap(~Location, scales = "free_x")

ggplot(dietsmelt, aes(x = smeltID, y = Contents.Weight))+ geom_col()+
  facet_wrap(Season~Location, scales = "free_x")

ggplot(dietsmelt, aes(x = Location, y = Contents.Weight))+ geom_boxplot()+
  facet_wrap(~Season, scales = "free_x")

ggplot(dietsmelt, aes(x = Location, y = Fullness))+ geom_boxplot()+
  facet_wrap(~Season, scales = "free_x")

ggplot(dietsmelt, aes(x = CageID2, y = Contents.Weight))+ geom_boxplot()+
  facet_wrap(Location~Season, scales = "free_x")


```

Can I statistically compare stomach fullness or weights between cage deployments?

Differences between locations in the fall will be the most useful, potentially.


```{r}
falldiet = filter(dietsmelt, Season == "Fall")
sf1 = lmer(Fullness ~ Location + (1|CageID2), data = falldiet)
summary(sf1)
plot(sf1)

library(emmeans)
plot(emmeans(sf1, pairwise ~ Location), comparison = T)
```

Nope. 

```{r}

fixef(sf1)["LocationSM"] = 0.1
fixef(sf1)["LocationYB"] = 0.1
#p1=powerSim(sf1, fixed("Location", "lr"))

#i'm getting nasty erros for all of teh power analyses with random effec tof cage and i on't know why


sf2 = lm(Fullness ~ Location, data = falldiet)
summary(sf2)

plot(sf2)

sf2$effects["LocationSM"] = 2

powerSim(sf2, progress = F)
#so, very, very low power

#how much power do we need?


model4 <- extend(sf2, within="Location", n=200)

#this takes forever to run, i'll do it tonight
pc4 <- powerCurve(model4,within="Location", breaks=c(50,100,150,200,250), progress =F)
print(pc4)

plot(pc4)
#so, no.

```

Now let's try diet contents weight

```{r}

hist(log(falldiet$Contents.Weight))

#convert weight to mg and log-transform
falldiet = mutate(falldiet, Weight = Contents.Weight*1000, logweight = log(Weight+1))

dw2 = lmer(logweight ~ Location + (1|CageID2), data = falldiet)
summary(dw2)
plot(emmeans(dw2, pairwise ~ Location), comparison = T)
plot(dw2)

fixef(dw2)["LocationYB"] = 0.2
#I don't know why i'm getting a bunch of 'is singular' errors, but we're still getting the power data
powerSim(dw2, fixed("LocationYB", "z"), progress = F)

#I'll try a simpler model without the random effect 
dw1 = lm(logweight ~ Location, data = falldiet)
summary(dw1)
plot(emmeans(dw1, pairwise ~ Location), comparison = T)
plot(dw1)

dw1$effects["LocationSM"] = .2

powerSim(dw1, progress =F)
#OK, that's better

#how much power do we need?


moddw <- extend(dw1, within="Location", n=200)

#power curve
pc5 <- powerCurve(moddw,within="Location", breaks=c(10, 20,30,40,50,100), progress =F)
print(pc5)

plot(pc5)
#so, we may be able to say something there.

```

Now let's look at growth data

```{r}
#source("smelt_2019_allseasons.Rmd")
#smeltdata2
load("Smeltdata.RData")
box_FL
box_K
box_Wt
```

Compare post-deployment weights in the fall

```{r}
fallsmelt = filter(smeltdata2, Deployment == "Fall", Period == "Post")

hist(fallsmelt$FL_cm)

#were lengths significantly different?
sml1 = lmer(FL_cm ~ Site  + (1|Cage), data = fallsmelt)
summary(sml1)
emmeans(sml1, pairwise~Site)


fixef(sml1)["SiteSM"] = 0.1
fixef(sml1)["SiteYolo"] = 0.1
fixef(sml1)["SiteFCCL"] = 0.1

powerSim(sml1, progress = F)


#check various sample sizes
#how much power do we need?


modfl <- extend(sml1, within="Site", n=200)

#power curve
pc6 <- powerCurve(modfl,within="Site", breaks=c(10, 20,30,40,50,100), progress =F)
print(pc6)

plot(pc6)
#so, we may be able to say something there.
```


```{r}
#what about condition factor?
hist(fallsmelt$K)

sml2 = lmer(K ~ Site  + (1|Cage), data = fallsmelt)
summary(sml2)
emmeans(sml2, pairwise~Site)

#Meh

fixef(sml2)["SiteSM"] = 0.02
fixef(sml2)["SiteYolo"] = 0.02
fixef(sml2)["SiteFCCL"] = 0.02

powerSim(sml2, progress = F)


#check various sample sizes
#how much power do we need?


modk <- extend(sml2, within="Site", n=200)

#power curve
pc6 <- powerCurve(modk,within="Site", breaks=c(10, 20,30,40,50,100), progress =F)
print(pc6)

plot(pc6)
#so, we may be able to say something there.


```



