---
title: "weight_analyses"
author: "Catarina Pien"
date: '2022-09-30'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls(all=TRUE))

library(car)
library(FSA)
library(tidyverse)
library(ggthemes)
library(here)
library(dplyr)
```

# Read in data
```{r}
# Add change in weight
Cagetype_data <- read.csv(here("smelt_2019_winterspring", "data_clean", "Cagetype_data.csv")) %>%
  mutate(deltaWt = Post_Weight_g - Pre_Weight_g,
         changeWt = round(deltaWt/Pre_Weight_g,2),
        deltaFL = Post_FL_cm - Pre_FL_cm,
         changeFL = round(deltaFL/Pre_FL_cm,2),
        Site = factor(Site, levels = c("RV", "DWSC")),
         Enclosure = ifelse(Mesh == "wrap", "A", ifelse(Mesh == "large", "B", ifelse(Mesh == "small", "C", ifelse(Mesh == "control", "control", NA)))))

# Make complete cases from column 9
Delta_CF <- Cagetype_data$Delta_CF
Cagetype_CF <- Cagetype_data[complete.cases(Delta_CF),]
Cageonly <- Cagetype_data %>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2") %>%
  mutate(Group = ifelse(Site == "RV" & Mesh == "wrap", "B", ifelse(Site == "DWSC", " ", "A")))

```

# Make manuscript boxplot
```{r}
windowsFonts(Times = windowsFont("Times New Roman"))

(Boxplot <- ggplot(data=Cageonly) +
   geom_boxplot(aes(x=Enclosure, y=deltaWt,  fill = Enclosure)) +
   geom_hline(yintercept = 0, size = 0.5) +
   facet_wrap(~Site) +
   # geom_text(aes(label = Group, x = Mesh, y = -0.32), family = "Times", size = 6) +
   scale_fill_brewer(palette="Dark2")+
   labs(y = "\u0394 Weight", x = "Enclosure Type") +
   theme_minimal() +
   annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
   annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
   theme(text = element_text(family = "Times"),
         axis.text = element_text(size = 16),
         axis.title = element_text(size = 17),
         strip.text = element_text(size = 17),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         panel.border = element_blank(),
         axis.line = element_line(colour = "black"),
         legend.position = "none",
         panel.spacing = unit(2, "lines")))

(Boxplot2 <- ggplot(data=Cageonly) +
    geom_boxplot(aes(x=Enclosure, y=changeWt,  fill = Enclosure)) +
    geom_hline(yintercept = 0, size = 0.5) +
    facet_wrap(~Site) +
    # geom_text(aes(label = Group, x = Mesh, y = -0.32), family = "Times", size = 6) +
    scale_fill_brewer(palette="Dark2")+
    labs(y = "Proportional Change in Weight", x = "Enclosure Type", labs = "Proportional Change in Weight") +
    theme_minimal() +
    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
    annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
    theme(text = element_text(family = "Times"),
          axis.text = element_text(size = 16),
          axis.title = element_text(size = 17),
          strip.text = element_text(size = 17),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = "none",
          panel.spacing = unit(2, "lines")))
```

Export figure
```{r}
tiff(filename=file.path("smelt_2019_winterspring/figures/Fig4_WeightByCage.tiff"), units="in",type="cairo", bg="white", height=4, 
     width=6, res=300, pointsize=12,compression="lzw")
Boxplot2
dev.off()
```

# Run analyses

## Pre-Post changes in weight and length
### RVR
```{r}
Cage_only_RVR <- Cagetype_CF %>%
  filter(Site == "RV")%>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2")
```

Mean
```{r}
Cage_only_RVR %>%
  summarize(meanWtPre = mean(Pre_Weight_g),
            meanWtPost = mean(Post_Weight_g),
            meandeltaWt = mean(deltaWt),
            n = n(),
            sddeltaWt = sd(deltaWt),
            sedeltaWt = sddeltaWt/n,
            meanchangeWt = mean(changeWt*100),
            sdchangeWt = sd(changeWt),
            sechangeWt = sdchangeWt/n,
            meanchangeFL =mean(changeFL*100),
            sechangeFL = sd(changeFL)/n,
            meanCFPre = mean(Pre_CF),
            meanCFPost = mean(Post_CF),
            seCF = sd(Pre_CF)/n,
            meandeltaCF = mean(Delta_CF),
            sedeltaCF = sd(Delta_CF)/n)
```

Does weight change in RVR? Yes
```{r}
hist(Cage_only_RVR$Pre_Weight_g)
hist(Cage_only_RVR$changeWt)
hist(log(Cage_only_RVR$changeWt+1))
hist(Cage_only_RVR$Post_Weight_g)
t.test(log(Cage_only_RVR$Pre_Weight_g+1), log(Cage_only_RVR$Post_Weight_g)+1)
```

Does length change in RVR? Yes
```{r}
hist(Cage_only_RVR$Pre_FL_cm)
hist(Cage_only_RVR$Post_FL_cm)
t.test(Cage_only_RVR$Pre_FL_cm, Cage_only_RVR$Post_FL_cm)
```

Does CF change in RVR?
```{r}
hist(Cage_only_RVR$Pre_CF)
hist(Cage_only_RVR$Post_CF)
t.test(log(Cage_only_RVR$Pre_CF+1), log(Cage_only_RVR$Post_CF+1))
```



### DWSC
```{r}
Cage_only_DWSC <- Cagetype_CF %>%
  filter(Site == "DWSC") %>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2")
```

Summary Stats
```{r}
Cage_only_DWSC %>%
  summarize(meanWtPre = mean(Pre_Weight_g),
            meanWtPost = mean(Post_Weight_g),
            meandeltaWt = mean(deltaWt),
            n = n(),
            sddeltaWt = sd(deltaWt),
            sedeltaWt = sddeltaWt/n,
            meanchangeWt = mean(changeWt*100),
            sdchangeWt = sd(changeWt),
            sechangeWt = sdchangeWt/n,
            meanchangeFL =mean(changeFL*100),
            sechangeFL = sd(changeFL)/n,
            meanCFPre = mean(Pre_CF),
            meanCFPost = mean(Post_CF),
            seCF = sd(Pre_CF)/n,
            meandeltaCF = mean(Delta_CF),
            sedeltaCF = sd(Delta_CF)/n)
```

T-test: Does weight change in DWSC? Yes
```{r}
hist(Cage_only_DWSC$Pre_Weight_g)
hist(Cage_only_DWSC$Post_Weight_g)
hist(Cage_only_DWSC$changeWt)
hist(log(Cage_only_DWSC$changeWt+1))
t.test(log(Cage_only_DWSC$Pre_Weight_g+1), log(Cage_only_DWSC$Post_Weight_g)+1)
```

Does length change in DWSC? No
```{r}
hist(Cage_only_DWSC$Pre_FL_cm)
hist(Cage_only_DWSC$Post_FL_cm)
t.test(Cage_only_DWSC$Pre_FL_cm, Cage_only_DWSC$Post_FL_cm)
```

Does CF change in DWSC? Yes
```{r}
t.test(log(Cage_only_DWSC$Pre_CF+1), log(Cage_only_DWSC$Post_CF)+1)
```

## Weight differences by cage type
### RVR

```{r}
library(lme4)
library(effects)
library(lmerTest)

# Checking if fish of each cage started at the same length
m_wtstart <- lm(Pre_Weight_g ~ Cage, data = Cage_only_RVR)
summary(m_wtstart)
anova((m_wtstart))
plot(m_wtstart)

# Simple LM
RV1 <- lm(changeWt ~ Enclosure, data = Cage_only_RVR)
summary(RV1)
anova(RV1)
# Residual standard dev = 0.117

# LMER - used this in manuscript
RV2 <- lmer(log(changeWt+1) ~ Mesh + (1|Cage), data = Cage_only_RVR, REML = TRUE)
summary(RV2)
anova(RV2)
plot(RV2)
qqnorm(resid(RV2))
# Residual standard dev = 0.113

confint(RV2)

```

```{r}
rv_residuals <- residuals(object = RV2)

par(mfrow = c(2,2))
plot(RV1)
leveneTest(log(changeWt+1) ~ Enclosure, data = Cage_only_RVR) # p<0.05 means not homogeneous; however, not too deviant in this case

shapiro.test(x = rv_residuals)

tab_model(RV2)
```


### DWSC


```{r}

# Checking if fish of each cage started at the same length
m_wtstart2 <- lm(Pre_Weight_g ~ Cage, data = Cage_only_DWSC)
summary(m_wtstart2)
anova((m_wtstart2))
plot(m_wtstart2)

# LM
DWSC1 <- lm(log(changeWt+1) ~ Enclosure, data = Cage_only_DWSC)
summary(DWSC1)
anova(DWSC1)
# Residual standard dev = 0.1675

# LMER
DWSC2 <- lmer(changeWt ~ Mesh + (1|Cage), data = Cage_only_DWSC, REML = TRUE)
summary(DWSC2)
anova(DWSC2)
# Residual standard dev = 0.1675

```

Check assumptions
```{r}
dwsc_resid= resid(DWSC1,type = "pearson")
par(mfrow = c(2,2))
plot(DWSC1)

leveneTest(DWSC1)
shapiro.test(log(Cage_only_DWSC$changeWt+1))
```

```{r}
library(sjPlot)
library(sjmisc)
library(sjlabelled)

tab_model(DWSC1)
```


## Length-Weight-CF Figures
```{r}
# Summarize mean, sd, median, n for Pre and Post Length and Width
Cagedata <- Cagetype_CF %>%
  na.omit() %>%
  filter(Cage!="FCCL1") %>%
  filter(Cage!="FCCL2") %>%
  group_by(Site, Mesh)%>%
  summarize(
    PreL_Mean = mean(Pre_FL_cm),
    PreL_sd = sd(Pre_FL_cm),
    PreL_median = median(Pre_FL_cm),
    PoL_Mean = mean(Post_FL_cm),
    PoL_sd = sd(Post_FL_cm),
    PoL_median = median(Post_FL_cm), 
    PreW_Mean = mean(Pre_Weight_g),
    PreW_sd = sd(Pre_Weight_g),
    PreW_median = median(Pre_Weight_g),
    PoW_Mean = mean(Post_Weight_g),
    PoW_sd = sd(Post_Weight_g),
    PoW_median = median(Post_Weight_g),  N = n()) %>%
    mutate(delta.Length = round(PoL_Mean - PreL_Mean,2)) %>%
  mutate(delta.weight = round(PoW_Mean - PreW_Mean,2)) %>%
  ungroup()
```

```{r}
Cageonly2 <- Cagetype_data %>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2") 

Cageonly <- Cagetype_data %>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2") %>%
  mutate(Group = ifelse(Site == "RV" & Mesh == "wrap", "B", ifelse(Site == "DWSC", " ", "A")))  %>%
  mutate(delta.L = round(Post_FL_cm - Pre_FL_cm,2))%>%
  mutate(delta.W = round(Post_Weight_g - Pre_Weight_g,2))



windowsFonts(Times = windowsFont("Times New Roman"))
```

Make length plot
```{r}
CagePrePost <- gather(Cageonly2, PrePost,Length1,Pre_FL_cm, Post_FL_cm)
CagePrePost$Site <- factor(CagePrePost$Site, levels = c("RV", "DWSC"))

MeanLengthplot<- ggplot(data=CagePrePost) +
  geom_boxplot(outlier.size=-1,aes(x=Enclosure, y=Length1,  fill = PrePost)) +
  facet_wrap(~Site) +
  labs(y = "Length (cm)") +
  #geom_text(aes(label = Group, x = Mesh, y = -0.32), family = "Times", size = 6) +
  #geom_hline(data = FCCL, aes(yintercept=delta.weight), size = 1) +
  #geom_text(data = ann_text, mapping = aes(x = x, y = y, label = label), family = "Times", size = 5.5)++
  scale_fill_manual(values=c("#E7298A", "#66A61E"), name = "", labels = c("Pre-Deployment", "Post-Deployment")) +
  labs(y = "Length (cm)", x = "Enclosure Type") + scale_x_discrete(labels=c("control" = "FCCL")) +
  theme_minimal() + theme(axis.title.x = element_blank()) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  #scale_fill_discrete(name = "", labels = c("Pre-Deployment", "Post-Deployment")) +
  theme(text = element_text(family = "Times"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 17),
        strip.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        #legend.position = "none",
        panel.spacing = unit(2, "lines"))
MeanLengthplot
```

Make weight plot
```{r}
CagePrePostW <- gather(Cageonly2, PrePostW,Weight1,Pre_Weight_g, Post_Weight_g)
windowsFonts(Times = windowsFont("Times New Roman"))

CagePrePostW$PrePostW <- factor(CagePrePostW$PrePostW,
                              levels = c('Pre_Weight_g','Post_Weight_g'),ordered = TRUE)
CagePrePostW$Site <- factor(CagePrePostW$Site, levels = c("RV", "DWSC"))

MeanWeightplot<- ggplot(data=CagePrePostW) +
  geom_boxplot(outlier.size=-1,aes(x=Enclosure, y=Weight1,  fill = PrePostW)) +
  facet_wrap(~Site) +
  labs(y = "Mean Weight (g)") +
  #geom_text(aes(label = Group, x = Mesh, y = -0.32), family = "Times", size = 6) +
  #geom_hline(data = FCCL, aes(yintercept=delta.weight), size = 1) +
  #geom_text(data = ann_text, mapping = aes(x = x, y = y, label = label), family = "Times", size = 5.5)+
  scale_fill_manual(values=c("#E7298A", "#66A61E"), name = "Deployment", labels = c("Pre-Deployment", "Post-Deployment"))+
  #palette="Dark2" 
   labs(y = "Weight (g)", x = "Enclosure Type") + scale_x_discrete(labels=c("control" = "FCCL")) +
  theme_minimal() + theme(strip.text.x = element_blank()) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) + 
  #scale_fill_discrete(name = "Deployment", labels = c("Pre-Deployment", "Post-Deployment")) +
  theme(text = element_text(family = "Times"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16),
        axis.text = element_text(size = 16),
        strip.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        #legend.position = "none",
        panel.spacing = unit(2, "lines"))
MeanWeightplot
```

Make CF plot

```{r}
CagePrePostK <- gather(Cageonly2, PrePostK,K1,Pre_CF, Post_CF)
windowsFonts(Times = windowsFont("Times New Roman"))

CagePrePostK$PrePostK <- factor(CagePrePostK$PrePostK,
                              levels = c('Pre_CF','Post_CF'),ordered = TRUE)
CagePrePostK$Site <- factor(CagePrePostK$Site, levels = c("RV", "DWSC"))

MeanKplot<- ggplot(data=CagePrePostK) +
  geom_boxplot(outlier.size=-1,aes(x=Enclosure, y=K1,  fill = PrePostK)) +
  facet_wrap(~Site) +
  labs(y = "Mean K") +
  #geom_text(aes(label = Group, x = Mesh, y = -0.32), family = "Times", size = 6) +
  #geom_hline(data = FCCL, aes(yintercept=delta.weight), size = 1) +
  #geom_text(data = ann_text, mapping = aes(x = x, y = y, label = label), family = "Times", size = 5.5)+
  scale_fill_manual(values=c("#E7298A", "#66A61E"), name = "Deployment", labels = c("Pre-Deployment", "Post-Deployment"))+
  #palette="Dark2" 
   labs(y = expression(K~(g~cm^-3)), x = "Enclosure Type") + scale_x_discrete(labels=c("control" = "FCCL")) +
  theme_minimal() + theme(strip.text.x = element_blank()) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) + 
  #scale_fill_discrete(name = "Deployment", labels = c("Pre-Deployment", "Post-Deployment")) +
  theme(text = element_text(family = "Times"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 17),
        strip.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        #legend.position = "none",
        panel.spacing = unit(2, "lines"))
MeanKplot
```

```{r}
library(patchwork)

lwkplot <- MeanLengthplot+MeanWeightplot+MeanKplot + plot_layout(nrow = 3, guides = "collect") &
  theme(legend.position = "bottom")
```

Export
```{r}
tiff(filename=file.path("smelt_2019_winterspring/figures/WeightLengthKPlot.tiff"), units="in",type="cairo", bg="white", height=7, 
     width=7, res=300, pointsize=12,compression="lzw")
lwkplot
dev.off()
```

## Survival

### RVR
Make data for RVR
```{r}
# Reorganize data
# Make complete cases from column 7 (post_weight)
post_fish <- Cagetype_data$Post_Weight_g
Cagetype_s <- Cagetype_data[complete.cases(post_fish),]

# ----------------Rio Vista-----------------------------------------
# Create data frame of each individual fish - survived or did not survive
Cage_survival_RV <- as.data.frame(matrix(0, ncol = 2, nrow = 384))
colnames(Cage_survival_RV) <- c("Cage", "Survived")

# Add dead fish from data sheet
  # Add RN82 from cage A that was decomposed
  # RG49 from cage A already included in the dead count (no Weight data)
  # Add RJ08 from cage E that was dead prior, stomach torn open

Cage_survival_RV$Cage <- c(rep("A", 64), rep("B", 64), rep("C", 64), rep("D", 64), rep("E", 64), rep("F", 64))
Cage_survival_RV$Survived <- c(rep("1", 60), rep("0", 4),
                               rep("1", 64), rep("0", 0),
                               rep("1", 64), rep("0", 0),
                               rep("1", 64), rep("0", 0),
                               rep("1", 62), rep("0", 2),
                               rep("1", 56), rep("0", 8))

Cagetype_survival_RV <- left_join(Cage_survival_RV, surv_RV, by = "Cage")

survival_RV <- Cagetype_survival_RV %>%
   mutate(Cage = factor(Cage),
         fSurvived = factor(Survived),
         Survived = as.numeric(Survived),
         Site = factor(Site),
         Mesh = factor(Mesh))%>%
  group_by(Mesh, Cage) %>%
  summarize(Survived = sum(Survived),
            n= n()) %>%
  mutate(rate = Survived/n*100)
```

Survival rate
```{r}
surv_RV %>%
  ungroup() %>%
  filter(Cage != "FCCL1") %>%
  summarize(meanRate = mean(rate),
            n = n(),
            seRate = sd(rate)/n)
```



Logistic Mixed Model
```{r}
mRV_surv <- glmer(Survived ~ Mesh + (1|Cage), data = Cagetype_survival_RV, family = binomial(link = "logit"))

summary(mRV_surv)
se <- sqrt(diag(vcov(mRV_surv)))
# table of estimates with 95% CI
(tab <- cbind(Est = fixef(mRV_surv), LL = fixef(mRV_surv) - 1.96 * se, UL = fixef(mRV_surv) + 1.96 *
                se))
```


### DWSC
```{r}
# Add dead fish
  # Add NTF4 to dead in Cage F - "decomposed"
  # Add RB61 to dead in Cage F - "partly decomposed"

Cage_survival_DWSC <- as.data.frame(matrix(0, ncol = 2, nrow = 360))
colnames(Cage_survival_DWSC) <- c("Cage", "Survived")

Cage_survival_DWSC$Cage <- c(rep("A", 60), rep("B", 60), rep("C", 60), rep("D", 60), rep("E", 60), rep("F", 60))
Cage_survival_DWSC$Survived <- c(rep("1", 59), rep("0", 1),
                                 rep("1", 60), rep("0", 0),
                                 rep("1", 58), rep("0", 2),
                                 rep("1", 60), rep("0", 0),
                                 rep("1", 60), rep("0", 0),
                                 rep("1", 58), rep("0", 2))

Cagetype_survival_DWSC <- left_join(Cage_survival_DWSC, surv_DWSC, by = "Cage")

survival_DWSC <- Cagetype_survival_DWSC %>%
  mutate(Cage = factor(Cage),
         fSurvived = factor(Survived),
         Survived = as.numeric(Survived),
         Site = factor(Site),
         Mesh = factor(Mesh))%>%
  group_by(Mesh, Cage) %>%
  summarize(Survived = sum(Survived),
            n= n()) %>%
  mutate(rate = Survived/n*100)
```

EDI table for survival
```{r}
survival_bind <- rbind(survival_DWSC, survival_RV) %>%
  mutate(rate = round(rate,2))
write_csv(survival_bind, here::here("smelt_2019_winterspring", "data_clean", "2019_smeltstudy_survival_data.csv"))
```


Survival rate
```{r}
Cagetype_survival_DWSC %>%
  select(-Survived) %>%
  distinct() %>%
  summarize(meanRate = mean(rate),
            n = n(),
            seRate = sd(rate)/n)
```



```{r}
mDWSC_surv <- glmer(Survived ~ Mesh + (1|Cage), data = Cagetype_survival_DWSC, family = binomial(link = "logit"))
# is singular

summary(mDWSC_surv)

se <- sqrt(diag(vcov(mDWSC_surv)))
# table of estimates with 95% CI
(tab <- cbind(Est = fixef(mDWSC_surv), LL = fixef(mDWSC_surv) - 1.96 * se, UL = fixef(mDWSC_surv) + 1.96 *
                se))
plot(mDWSC_surv)

r.squaredGLMM(mDWSC_surv)
plot_model(mDWSC_surv, type = "pred", terms = c("Mesh"))
tab_model(mDWSC_surv)
```






### Plots
```{r}
# Survival Plot ---------------------------------------
# * By Mesh ----
RV_survival_sum <- Cagetype_survival_RV %>%
  group_by(Mesh) %>%
  summarize(Survive = sum(Survived == 1),
            Died = -1 * sum(Survived == 0),
            n = n()) %>%
  pivot_longer(cols = c(Survive, Died), names_to = "Survival", values_to = "Count")

DWSC_survival_sum <- Cagetype_survival_DWSC %>%
  group_by(Mesh) %>%
  summarize(Survive = sum(Survived == 1),
            Died = -1 * sum(Survived == 0),
            n = n()) %>%
  pivot_longer(cols = c(Survive, Died), names_to = "Survival", values_to = "Count")

DWSC_survival_sum$Site <- "DWSC"
RV_survival_sum$Site <- "RV"

Survival_sum <- rbind(DWSC_survival_sum, RV_survival_sum)
Survival_sum <- Survival_sum %>%
  mutate(prop = ifelse(Survival == "Survive", round(Count/n * 100,1), round((-1 * Count/n * 100),1)))

# Survival plot
ggplot(Survival_sum, aes(Mesh, Count, fill = Survival)) +
  facet_wrap(~Site) +
  geom_col(width = 0.55) +
  coord_cartesian(y = c(-30,130)) +
  scale_y_continuous(breaks = seq(0, 125, by = 25)) +
  geom_text(aes(label = paste0(prop, "%")), vjust = 1.5, size = 5)+
  geom_hline(yintercept=0)+
  scale_fill_brewer(palette="Set1")+
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title = element_text(size = 17),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 16),
        legend.position = "none")



# * By Cage ----
RV_survival_sum2 <- Cagetype_survival_RV %>%
  group_by(Mesh, Cage) %>%
  summarize(Survive = sum(Survived == 1),
            Died = -1 * sum(Survived == 0),
            n = n()) %>%
  pivot_longer(cols = c(Survive, Died), names_to = "Survival", values_to = "Count")

DWSC_survival_sum2 <- Cagetype_survival_DWSC %>%
  group_by(Mesh, Cage) %>%
  summarize(Survive = sum(Survived == 1),
            Died = -1 * sum(Survived == 0),
            n = n()) %>%
  pivot_longer(cols = c(Survive, Died), names_to = "Survival", values_to = "Count")

DWSC_survival_sum2$Site <- "DWSC"
RV_survival_sum2$Site <- "RV"

Survival_sum2 <- rbind(DWSC_survival_sum2, RV_survival_sum2)
Survival_sum2 <- Survival_sum2 %>%
  mutate(prop = ifelse(Survival == "Survive", round(Count/n * 100,1), round((-1 * Count/n * 100),1)))

ggplot(Survival_sum2, aes(Cage, Count, fill = Survival)) +
  facet_wrap(~Site) +
  geom_col(width = 0.7, aes(colour = Mesh), size = 3) +
  coord_cartesian(y = c(-30,130)) +
  scale_y_continuous(breaks = seq(0, 125, by = 25)) +
  scale_color_brewer(palette="Dark2")+
  geom_text(aes(label = paste0(prop, "%")), vjust = 1.5, size = 5)+
  geom_hline(yintercept=0)+
  scale_fill_brewer(palette="Set1")+
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title = element_text(size = 17),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 16),
        legend.position = "none")

### Mean survival

Survival_sum3 <- Survival_sum2 %>%
  mutate(prop = ifelse(Survival == "Survive", Count/n * 100, -1 * Count/n * 100))

propSurvival <- filter(Survival_sum3, Survival == "Survive") %>%
  ungroup() %>%
group_by(Site) %>%
  summarize(meanSurvival = mean(prop),
            n = n(),
            seSurvival = sd(prop)/n)

```






### Old Code
```{r}
# Logistic mixed model ----

library(lme4)
library(lmerTest)

mRV_surv <- glmer(Survived ~ Mesh + (1|Cage), data = Cagetype_survival_RV, family = binomial(link = "logit"))

summary(mRV_surv)
se <- sqrt(diag(vcov(mRV_surv)))
# table of estimates with 95% CI
(tab <- cbind(Est = fixef(mRV_surv), LL = fixef(mRV_surv) - 1.96 * se, UL = fixef(mRV_surv) + 1.96 *
                se))
# * * Check assumptions ----
Ms1.resid <- resid(mRV_surv, type = "pearson")
plot(mRV_surv)
plot(mRV_surv, form = resid(., type = "pearson") ~ fitted(.) | Mesh, abline = 0,
     cex = .5, pch = 20, col = "black")

# normality
par(mfrow = c(1,1))
qqnorm(Ms1.resid)
qqline(Ms1.resid)
hist(Ms1.resid)

# observed vs fitted
# There is no funnel
plot(mRV_surv, Survived ~ fitted(.), id = 0.05, adj = -0.3,
     cex = .8, pch = 20, col = "blue")
# leverage
ggplot(data.frame(lev=hatvalues(mRV_surv),pearson=residuals(mRV_surv,type="pearson")),
       aes(x=lev,y=pearson)) +
  geom_point() +
  theme_bw()

# variance by mesh
ggplot(Cagetype_survival_RV, aes(x = Mesh, y = Ms1.resid)) + geom_boxplot()

leveneTest(Survived ~ Mesh, data = Cage_only_DWSC)

# * * Plot results ----
library(MuMIn) #extract R2
library(sjPlot)
library(sjstats) # for intraclass-correlation coefficient
library(caret) # model comparison & utility functions

r.squaredGLMM(mRV_surv)
plot_model(mRV_surv, type = "pred", terms = c("Mesh"))
tab_model(mRV_surv)

library(ggeffects)
ggpredict(mRV_surv, "Survived")





# Logistic mixed model ----

library(lme4)
library(lmerTest)

#logit transform survival data
#lmer with random effect cage
#anova

mDWSC_surv <- glmer(Survived ~ Mesh + (1|Cage), data = Cagetype_survival_DWSC, family = binomial(link = "logit"))
# is singular

summary(mDWSC_surv)

se <- sqrt(diag(vcov(mDWSC_surv)))
# table of estimates with 95% CI
(tab <- cbind(Est = fixef(mDWSC_surv), LL = fixef(mDWSC_surv) - 1.96 * se, UL = fixef(mDWSC_surv) + 1.96 *
                se))
plot(mDWSC_surv)

# * * Check assumptions ----
Ms2.resid <- resid(mDWSC_surv, type = "pearson")
plot(mDWSC_surv)
plot(mDWSC_surv, form = resid(., type = "pearson") ~ fitted(.) | Mesh, abline = 0,
     cex = .5, pch = 20, col = "black")

# normality
par(mfrow = c(1,1))
qqnorm(Ms2.resid)
qqline(Ms2.resid)
hist(Ms2.resid)

# leverage
ggplot(data.frame(lev=hatvalues(mDWSC_surv),pearson=residuals(mDWSC_surv,type="pearson")),
       aes(x=lev,y=pearson)) +
  geom_point() +
  theme_bw()

# check variance
ggplot(Cagetype_survival_DWSC, aes(x = Mesh, y = Ms2.resid)) + geom_boxplot()


# * * Plot results ----
library(MuMIn) #extract R2
library(sjPlot)
library(sjstats) # for intraclass-correlation coefficient
library(caret) # model comparison & utility functions







