---
title: "Relative Condition Factor"
author: "Catarina Pien"
date: "8/31/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls(all=TRUE))
library(tidyverse)
library(here)

```

Read in data

```{r import}
Cagetype_data <- read.csv(here("smelt_2019_winterspring", "data_clean", "Cagetype_data.csv"))
Cagetype_data$Site <- ordered(Cagetype_data$Site, levels = c("RV", "DWSC"))
```

Complete cases

```{r complete, echo=FALSE}
# Make complete cases from column 9
Delta_CF <- Cagetype_data[,9]
Cagetype_CF <- Cagetype_data[complete.cases(Delta_CF),]
```

Gather length weight data and log transform
```{r}
LWData <- Cagetype_CF %>%
 select(-contains("CF")) %>%
  rename(FL_Pre = Pre_FL_cm,
         FL_Post = Post_FL_cm,
         Weight_Pre = Pre_Weight_g,
         Weight_Post = Post_Weight_g) %>%
  mutate(Site2 = ifelse(Mesh == "control", "FCCL", as.character(Site)))
         
# .value: this tells pivot_longer() that that part of the column name specifies the “value” being measured (which will become a variable in the output)
LWlong <- LWData %>% 
  pivot_longer(cols = -c(X, Cage, Site2, Site, Mesh),
               names_to = c(".value", "Event"),
               names_sep = "_") 

LWlongT <- LWlong %>%
  mutate(logFL = log(FL/10),
         logWeight = log(Weight))

```

Plot data
```{r plot}
# Overall
ggplot(LWlongT, aes(x = FL, y = Weight)) + geom_point() + 
  labs(x = "Fork Length (cm)",
       y = "Weight (g)") + 
  theme_bw()

# Site
ggplot(LWlongT, aes(x = logFL, y = logWeight, shape = Site2, color = Site)) + geom_point() + 
  geom_smooth(aes(color = Site), method = "lm", se = TRUE)+
  scale_shape_manual(values = c(19, 2))+
  labs(x = "Log10 Fork Length (mm)",
       y = "Log10 Weight (g)") + 
  theme_bw()

# Pre-Post
ggplot(LWlongT, aes(x = logFL, y = logWeight, shape = Event, color = Event)) + geom_point() + 
  geom_smooth(aes(color = Event), method = "lm", se = TRUE)+
  facet_wrap(~Site2) + 
  labs(x = "Log10 Fork Length (mm)",
       y = "Log10 Weight (g)") + 
  theme_bw()
```

Fit a model to the LW-data
```{r glm}
library(visreg)
library(lsmeans)
# Linear model  - all data
m1 <- lm(logWeight ~ logFL, data = LWlongT)
summary(m1)
par(mfrow = c(2,2))
plot(m1) #3.31
visreg(m1)

# RV only
RV_LW = filter(LWlongT, Site2 == "RV")
m2 <- lm(logWeight ~ logFL, data = RV_LW)
summary(m2)
plot(m2) #3.35
visreg(m2)

# DWSC only

DWSC_LW = filter(LWlongT, Site2 == "DWSC")
m3 <- lm(logWeight ~ logFL, data = DWSC_LW)
summary(m3) #3.51
plot(m3)
visreg(m3)

# DWSC only

FCCL_LW = filter(LWlongT, Site2 == "FCCL")
m4 <- lm(logWeight ~ logFL, data = FCCL_LW)
summary(m4) #3.62
plot(m4)
visreg(m4)


```

### Testing differences in L-W Relationship (ANCOVA)
```{r}
# Are sites different? Yes.
m5 <- lm(logWeight ~ logFL*Site2, data = LWlongT)
anova(m5)
m5$coefficients
(m.lst <- lstrends(m5, "Site2", var="logFL"))
pairs(m.lst)

# Plot Sites
ggplot(LWlongT, aes(x = logWeight, y = logFL, color = Site2)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) + 
  facet_wrap(~Event) + theme_bw()

# Are Pre-Post different?
m6 <- lm(logWeight ~ logFL*Event + Site2, data = LWlongT)
anova(m6)
m6$coefficients
(m.lst2 <- lstrends(m6, "Event", var="logFL")) # 3.387
pairs(m.lst2)

# Plot Event
ggplot(LWlongT, aes(x = logWeight, y = logFL, color = Event)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  
  facet_wrap(~Site2) + theme_bw()

# Including Both
m7 <- lm(logWeight ~ logFL*Event + logFL*Site2, data = LWlongT)
anova(m7)
m7$coefficients
(m.lst3 <- lstrends(m6, "Event", var="logFL")) # 3.387
pairs(m.lst3)

```

Condition Factor
LW regression: logW = 2.01266 + 3.30975 x logFL
Relative CF: Kn = W/(a x L^b)
Fulton's CF: W/L^3
Ricker's CF = W/L^b
```{r CF}
# a = 10^intercept
# b = slope of line

a = 102.958
b = 3.30975

CFData <- LWData %>%
  mutate(
    FL_Pre_mm = FL_Pre/10,
    FL_Post_mm = FL_Post/10,
    RelCF_Pre = Weight_Pre/(a * FL_Pre_mm^b) ,
    RelCF_Post = Weight_Post/(a * FL_Post_mm^b),
    Delta_RelCF = RelCF_Post-RelCF_Pre,
    RickCF_Pre = (Weight_Pre/(FL_Pre^b)) * 100,
    RickCF_Post = (Weight_Post/(FL_Post^b)) * 100,
    Delta_RickCF = RickCF_Post-RickCF_Pre,
    CF_Pre = Weight_Pre/(FL_Pre^3)*100,
    CF_Post = Weight_Post/(FL_Post^3)*100,
    Delta_CF = CF_Post-CF_Pre)
  

```

Mean CF for plots

```{r meancf Plots}
# Summarize mean, sd, median, n for condition factor
Cage.CF <- CFData %>%
  na.omit() %>%
#  filter(Cage!="FCCL1") %>%
 # filter(Cage!="FCCL2") %>%
  group_by(Site, Mesh)%>%
  summarize(
    Delta_CF_Mean = mean(Delta_CF),
    Delta_CF_sd = sd(Delta_CF),
    Delta_CF_median = median(Delta_CF),
    N = n(),
    Delta_RCF_Mean = mean(Delta_RickCF),
    Delta_RCF_sd = sd(Delta_RickCF),
    Delta_RCF_median = median(Delta_RickCF),
    NRick = n())

# Calculate Average CF for FCCL
FCCL_CF <- CFData %>%
  filter(Cage == "FCCL1" | Cage == "FCCL2") %>%
  group_by(Site, Mesh)%>%
  summarize(
    Delta_CF_Mean = mean(Delta_CF),
    Delta_CF_sd = sd(Delta_CF),
    Delta_CF_median = median(Delta_CF),
    N = n(),
    Delta_RCF_Mean = mean(Delta_RickCF),
    Delta_RCF_sd = sd(Delta_RickCF),
    Delta_RCF_median = median(Delta_RickCF),
    NRick = n()
  )
```

Plots
```{r plots}
# Overview of data (geom_point)
ggplot(CFData, aes(Mesh, Delta_CF, col = Mesh)) + 
  facet_wrap(~Site) +
  geom_jitter(size = 3) +
  scale_fill_brewer(palette="Dark2")+
  theme_bw() +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18))

# CF mean and sd by site
# CF mean by site
ggplot(Cage.CF, aes(Mesh, Delta_CF_Mean,  fill = Mesh)) + 
  facet_wrap(~Site)+
  geom_bar(width = 0.8, position = "dodge2", stat = "identity") +
  scale_fill_brewer(palette="Dark2")+
  labs(y = "Delta Condition Factor") +
  geom_errorbar(aes(ymin = Delta_CF_Mean - Delta_CF_sd, ymax = Delta_CF_Mean + Delta_CF_sd), width = 0.1, 
                position = position_dodge(0.9), color = "black") +
  theme_bw() +
  theme(axis.text = element_text(size = 18),
        axis.title = element_text(size = 18),
        strip.text = element_text(size = 18),
        legend.position = "none")

# Mean point plot
ggplot(Cage.CF, aes(Mesh, Delta_CF_Mean,  col = Mesh)) + 
  facet_wrap(~Site)+
  geom_point(size = 8, shape = 18) +
  scale_colour_brewer(palette="Dark2")+
  labs(y = "Delta Condition Factor") +
  geom_errorbar(aes(ymin = Delta_CF_Mean - Delta_CF_sd, ymax = Delta_CF_Mean + Delta_CF_sd), width = 0.1, 
                position = position_dodge(0.9), color = "black") +
  theme_bw() +
  theme(axis.text = element_text(size = 18),
        axis.title = element_text(size = 18),
        strip.text = element_text(size = 18),
        legend.position = "none")

# Mean point plot - Ricker
ggplot(Cage.CF, aes(Mesh, Delta_RCF_Mean,  col = Mesh)) + 
  facet_wrap(~Site)+
  geom_point(size = 8, shape = 18) +
  scale_colour_brewer(palette="Dark2")+
  labs(y = "Delta Condition Factor") +
  geom_errorbar(aes(ymin = Delta_RCF_Mean - Delta_RCF_sd, ymax = Delta_RCF_Mean + Delta_RCF_sd), width = 0.1, 
                position = position_dodge(0.9), color = "black") +
  theme_bw() +
  theme(axis.text = element_text(size = 18),
        axis.title = element_text(size = 18),
        strip.text = element_text(size = 18),
        legend.position = "none")

### CF boxplot by site #####################
Cageonly <- CFData %>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2") %>%
  mutate(Group = ifelse(Site == "RV" & Mesh == "wrap", "B", ifelse(Site == "DWSC", " ", "A")))

windowsFonts(Times = windowsFont("Times New Roman"))

# Specify FCCL Delta CF label position
ann_text <- data.frame(
  label = c("FCCL Delta CF = 0.0048","FCCL Delta CF = 0.0027"),
  Site = c("RV", "DWSC"),
  x = c(2,2),
  y = c(0.08, 0.07))

# By mesh type
ggplot(data=Cageonly) +
  geom_boxplot(aes(x=Mesh, y=Delta_CF,  fill = Mesh)) +
  facet_wrap(~Site) +
  labs(y = "Delta Condition Factor") +
  geom_text(aes(label = Group, x = Mesh, y = -0.05), family = "Times", size = 6) +
  geom_hline(data = FCCL_CF, aes(yintercept=Delta_CF_median), size = 1) +
  geom_text(data = ann_text, mapping = aes(x = x, y = y, label = label), family = "Times", size = 5.5)+
  scale_fill_brewer(palette="Dark2")+
  labs(y = "Delta Condition Factor") +
  theme_minimal() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  theme(text = element_text(family = "Times"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 17),
        strip.text = element_text(size = 17),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none",
        panel.spacing = unit(2, "lines"))

# Specify FCCL Delta CF label position
ann_text2 <- data.frame(
  label = c("FCCL Delta CF = 0.0135","FCCL Delta CF = 0.024"),
  Site = c("RV", "DWSC"),
  x = c(2,2),
  y = c(0.08, 0.07))

# By mesh type - Ricker
ggplot(data=Cageonly) +
  geom_boxplot(aes(x=Mesh, y=Delta_RelCF,  fill = Mesh)) +
  facet_wrap(~Site) +
  labs(y = "Delta Condition Factor") +
  geom_text(aes(label = Group, x = Mesh, y = -0.05), family = "Times", size = 6) +
  geom_hline(data = FCCL_CF, aes(yintercept=Delta_RCF_median), size = 1) +
  geom_text(data = ann_text2, mapping = aes(x = x, y = y, label = label), family = "Times", size = 5.5)+
  scale_fill_brewer(palette="Dark2")+
  labs(y = "Delta Condition Factor") +
  theme_minimal() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  theme(text = element_text(family = "Times"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 17),
        strip.text = element_text(size = 17),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none",
        panel.spacing = unit(2, "lines"))
# By cage
ggplot(data=Cageonly, aes(x=Cage, y=Delta_CF,  fill = Mesh)) +
  facet_wrap(~Site) +
  labs(y = "Delta Condition Factor") +
  geom_boxplot()+
  geom_hline(data = FCCL_CF, aes(yintercept=Delta_CF_median)) +
  scale_fill_brewer(palette="Dark2")+
  labs(y = "Delta Condition Factor") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 17),
        strip.text = element_text(size = 16),
        legend.position = "none")

```


ANOVA

```{r anova, warning = FALSE, message = FALSE}
library(car)
library(FSA)
## RV ---------------------------------------------------------------------------
# Run ANOVA -----------------------------

Cage_only_RVR <- CFData %>%
  filter(Site == "RV")%>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2")

rv.cf.aov <- aov(Delta_CF~Mesh, data = Cage_only_RVR)
summary(rv.cf.aov)

# Check assumptions -----------------------------------
# Homogeneity of variances
par(mfrow = c(1,2))
plot(rv.cf.aov, 1)
leveneTest(Delta_CF ~ Mesh, data = Cage_only_RVR) # p<0.05 means not homogeneous

# Normality
plot(rv.cf.aov, 2)

rv_cf_residuals <- residuals(object = rv.cf.aov)
shapiro.test(x = rv_cf_residuals ) # p<0.05 means not normal

# Does not meet assumptions: do non-parametric Kruskal-Wallis test--------------------
kruskal.test(Delta_CF~Mesh, data = Cage_only_RVR) # still significant

# Post-hoc Dunn test --------------
# Dunn test can be used if sample size are heterogeneous
RV.dunn <- dunnTest(Delta_CF~Mesh,
                    data = Cage_only_RVR,
                    method = "bonferroni") # bonferroni corrects for number of comparisons
RV.dunn
RV.dunn = RV.dunn$res

library(rcompanion)

cldList(comparison = RV.dunn$Comparison,
        p.value    = RV.dunn$P.adj,
        threshold  = 0.05)

## DWSC ---------------------------------------------------------------------------

# Run ANOVA -------------------------------------

Cage_only_DWSC <- CFData %>%
  filter(Site == "DWSC") %>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2")

Cage_only_DWSC$Mesh <- as.factor(Cage_only_DWSC$Mesh) 

dwsc.cf.aov <- aov(Delta_CF~Mesh, data = Cage_only_DWSC)
summary(dwsc.cf.aov)

TukeyHSD(dwsc.cf.aov)

# Check assumptions -----------------------------------
# Homogeneity of variances
plot(dwsc.cf.aov, 1)
leveneTest(Delta_CF ~ Mesh, data = Cage_only_DWSC) # p<0.05 means not homogeneous

# Normality
plot(dwsc.cf.aov, 2)

dwsc_cf_residuals <- residuals(object = dwsc.cf.aov)
shapiro.test(x = dwsc_cf_residuals ) # p<0.05 means not normal

# Does not meet assumptions: do non-parametric Kruskal-Wallis test---------------
kruskal.test(Delta_CF~Mesh, data = Cage_only_DWSC)
```

Ricker CF - ANOVA /Kruskal Wallis
```{r Rickeranova, warning = FALSE, message = FALSE}
library(car)
library(FSA)
## RV ---------------------------------------------------------------------------
# Run ANOVA -----------------------------

Cage_only_RVR <- CFData %>%
  filter(Site == "RV")%>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2")

rv.cf.aov <- aov(Delta_RickCF~Mesh, data = Cage_only_RVR)
summary(rv.cf.aov)

# Check assumptions -----------------------------------
# Homogeneity of variances
par(mfrow = c(1,2))
plot(rv.cf.aov, 1)
leveneTest(Delta_RickCF ~ Mesh, data = Cage_only_RVR) # p<0.05 means not homogeneous

# Normality
plot(rv.cf.aov, 2)

rv_cf_residuals <- residuals(object = rv.cf.aov)
shapiro.test(x = rv_cf_residuals ) # p<0.05 means not normal

# Does not meet assumptions: do non-parametric Kruskal-Wallis test--------------------
kruskal.test(Delta_RickCF~Mesh, data = Cage_only_RVR) # still significant

# Post-hoc Dunn test --------------
# Dunn test can be used if sample size are heterogeneous
RV.dunn <- dunnTest(Delta_RickCF~Mesh,
                    data = Cage_only_RVR,
                    method = "bonferroni") # bonferroni corrects for number of comparisons
RV.dunn
RV.dunn = RV.dunn$res

library(rcompanion)

cldList(comparison = RV.dunn$Comparison,
        p.value    = RV.dunn$P.adj,
        threshold  = 0.05)

## DWSC ---------------------------------------------------------------------------

# Run ANOVA -------------------------------------

Cage_only_DWSC <- CFData %>%
  filter(Site == "DWSC") %>%
  filter(Cage != "FCCL1") %>%
  filter(Cage != "FCCL2")

Cage_only_DWSC$Mesh <- as.factor(Cage_only_DWSC$Mesh) 

dwsc.cf.aov <- aov(Delta_RickCF~Mesh, data = Cage_only_DWSC)
summary(dwsc.cf.aov)

TukeyHSD(dwsc.cf.aov)

# Check assumptions -----------------------------------
# Homogeneity of variances
plot(dwsc.cf.aov, 1)
leveneTest(Delta_RickCF ~ Mesh, data = Cage_only_DWSC) # p<0.05 means not homogeneous

# Normality
plot(dwsc.cf.aov, 2)

dwsc_cf_residuals <- residuals(object = dwsc.cf.aov)
shapiro.test(x = dwsc_cf_residuals ) # p<0.05 means not normal

# Does not meet assumptions: do non-parametric Kruskal-Wallis test---------------
kruskal.test(Delta_RickCF~Mesh, data = Cage_only_DWSC)
```


# Mixed Modeling to incorporate the effect of cage? Reach out to Britt or Rosie for advice. 
```{r}
library(lme4)
library(lmerTest)

M.RVR = lmer(Delta_CF ~ Mesh + (1|Cage), data = Cage_only_RVR, na.action = na.omit)
summary(M.RVR)
anova(M.RVR)

M1.resid <- resid(M.RVR, type = "pearson")
plot(M.RVR)
par(mfrow = c(2,2))

# normality
qqnorm(residuals(M.RVR))
# leverage
ggplot(data.frame(lev=hatvalues(M.RVR),pearson=residuals(M.RVR,type="pearson")),
      aes(x=lev,y=pearson)) +
    geom_point() +
    theme_bw()

ggplot(Cage_only_RVR, aes(y = Mesh, x = M1.resid)) + geom_point()

plot(hist(residuals(M.RVR)))


M.RVR2 = lmer(Delta_CF ~ Mesh + (1|Cage), data = Cage_only_RVR, na.action = na.omit)
summary(M.RVR)
anova(M.RVR)

par(mfrow = c(1,2))
qqnorm(residuals(M.RVR))
plot(residuals(M.RVR))



M.DWSC = lmer(Delta_CF ~ Mesh + (1|Cage), data = Cage_only_DWSC, na.action = na.omit)
summary(M.DWSC)
qqnorm(residuals(M.DWSC))
plot(residuals(M.DWSC))

```

