---
title: "Recpature distances"
author: "Rosie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(tidyverse)
library(readxl)
library(sf)
library(lubridate)
library(spacetools)
library(deltamapr)
#load("Recaps.RData")
#save(nndist, distances, recaps, recaps2, recaps3, RVhdist, SCsoft, SChard, RVhdistx, Dwschard, Dwscsoft, file = "Recaps.RData")
```

## Spatial stats on Delta Smelt releases

We have four different release types and a few questions we can ask.

1. How far did smelt from each release group travel before being recaptured?
2. How variable was the distance between release and recapture?
3. How close together were fish from a particular release when they were recaptured?
4. How variable was the distance between fish on recapture?

Sam Bashevkin has written some nifty functions for calculating in=water distance between points (fish) and a reference point (release location) or between a group of points (fish from a particular release). 

https://github.com/sbashevkin/spacetools

Fist let's calculate distance from release location to recapture location.

```{r}
#Bring in the data and check it out


recaps = read_excel("data/Running Delta Smelt Catch.xlsx", 
                    sheet = "Delta Smelt Catch Data") %>%
  rename( Latitude= `LatitudeStart`, Longitude = `LongitudeStart`)

recaps = recaps %>%
  mutate(FishID = paste("Fish", 1:nrow(recaps))) %>%
  select(-ReleaseDate, -Preservation, -Disposition, -SpecialStudyID, -MarkCode, -Dateknown)


#bring in standaridized release location/mark info
release = read_excel("data/releases_v2.xlsx") %>%
  rename(LatitudeR = Latitude, LongitudeR = Longitude) 


recaps2 = left_join(recaps, select(release, -TotalReleased), by = "ReleaseEvent") %>%
  mutate(DaysSinceRelease = SampleDate - ReleaseDate,
         Type = case_when(ReleaseMethod == "Unknown" ~ "Unknown",
                          ReleaseMethod == "NA" ~ "Wild/unmarked",
                          TRUE ~ Type))

str(recaps)



```

This just needs a little data manipulation

```{r}
releasesf = st_as_sf(filter(release, !is.na(LatitudeR)),  coords = c("LongitudeR", "LatitudeR"), crs = 4326)

#make a spatial data frame of recaptures

recapsf = st_as_sf(filter(recaps2, !is.na(`Longitude`)), coords = c("Longitude", "Latitude"), crs = 4326)

#make a quick plot to make sure I did it right
ggplot()+
  geom_sf(data = WW_Delta)+
  geom_sf(data = recapsf, aes(color = Type))+
  geom_sf(data = releasesf, aes(color = Type), size = 5)+
  
  coord_sf(xlim = c(-121.5, -122.2), ylim = c(37.8, 38.5))


```


Now we can calculate the distance between each point and the release location. I'm going to do it one release at a time. THere is probalby a way to vectorize this and do it all at once, but I'm too lazy to figure it out. 

```{r, eval=FALSE}
#one release at a time
Dwschard = filter(recaps2, Type == "Hard (carboy)", Site == "SDWSC")
dwsch = filter(release, Site == "SDWSC", Type == "Hard (carboy)") %>%
  select(LatitudeR, LongitudeR) %>%
  rename(Latitude = LatitudeR, Longitude = LongitudeR)
tesst = GGdist(Water_map = spacetools::Delta, Points = Dwschard, Latitude_column = `Latitude`,
               Longitude_column = `Longitude`, PointID_column = FishID, EndPoint = dwsch)

#now let's do DWSC soft
Dwscsoft = filter(recaps2, Type == "Soft (carboy)", Site == "SDWSC")
dwscs = filter(release, Site == "SDWSC", Type == "Soft (carboy)") %>%
  select(LatitudeR, LongitudeR)%>%
  rename(Latitude = LatitudeR, Longitude = LongitudeR)
tesst2 = GGdist(Water_map = spacetools::Delta, Points = Dwscsoft, Latitude_column = `Latitude`,
               Longitude_column = `Longitude`, PointID_column = FishID, EndPoint = dwscs)

#now let's do Rio Vista hard
RVhard = filter(recaps2,  Site == "Rio Vista", !is.na(Latitude))
rvh = filter(release, Site == "Rio Vista", Type == "Hard (trailer)") %>%
  select(LatitudeR, LongitudeR)%>%
  rename(Latitude = LatitudeR, Longitude = LongitudeR) %>%
  distinct()
tesst3 = GGdist(Water_map = spacetools::Delta, Points = RVhard, Latitude_column = `Latitude`,
                Longitude_column = `Longitude`, PointID_column = FishID, EndPoint = rvh)

# #now the trailer
# RVhardt = filter(recaps2, Type == "Hard (Trailer)", Site == "Rio Vista")
# rvht = filter(release, Site == "Rio Vista", Type == "Hard (Trailer)") %>%
#   select(LatitudeR, LongitudeR)%>%
#   rename(Latitude = LatitudeR, Longitude = LongitudeR)
# tesst4 = GGdist(Water_map = spacetools::Delta, Points = RVhardt, Latitude_column = `Latitude`,
#                 Longitude_column = `Longitude`, PointID_column = fishID, EndPoint = rvht)

distances = bind_rows(tesst, tesst2, tesst3)
```

OK, now let's bind all the distances together and join them back to the origional dataset

```{r}

recaps3 = left_join(recaps2, distances) %>%
  mutate(Type2 = case_when(Type %in% c("Hard (large scale)","Hard (trailer)") ~ "Hard (Large Scale)",
                           TRUE ~ Type)) %>%
  mutate(SiteType = paste(Type, Site))

recaps3

#load first flush info
ff = read_excel("data/releases_v2.xlsx", sheet = "firstflush")

recaps3 = recaps3 %>%
  mutate(Year = year(SampleDate),
                                BroodYear = case_when(LifeStage %in% c("Adult", "Adult*") & yday(SampleDate) <200 ~ Year-1,
                                     LifeStage == "Juvenile" & yday(SampleDate) <90 ~ Year-1,
                                     TRUE ~ Year)) %>%
         left_join(ff)%>%
         mutate(FFyn = case_when(ReleaseDate > FirstFlush ~ "Yes",
                          TRUE ~ "No"))

write.csv(recaps, "recaptureanalysis/Recapturedistances.csv")
```

Plot to visually look at mean and variation.

```{r}
ggplot(filter(recaps3, Type != "Wild/unmarked", Type != "Unknown"), aes(x = SiteType, y = Distance))+ 
  geom_boxplot(aes(fill = SiteType))+
  ylab("Distance from Release Site (m)")+
  scale_fill_discrete(guide = NULL)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust =1))
```
now by first flush

```{r}

ggplot(filter(recaps3, Type != "Wild/unmarked", Type != "Unknown"), aes(x = Type, y = Distance))+ 
  geom_boxplot(aes(fill = SiteType))+
  ylab("Distance from Release Site (m)")+
  scale_fill_discrete(guide = NULL)+
  theme_bw()+
  facet_wrap(~FFyn)+
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust =1))

```


Let's consolidate some of the release types...

```{r}

tots = group_by(recaps3, Type2) %>%
  summarize(N = n())

ggplot(filter(recaps3, Type != "Wild/unmarked", Type != "Unknown"), aes(x = Type2, y = Distance/1000))+ 
  geom_boxplot(aes(fill = Type2))+
  ylab("Distance from Release Site (km)")+
  scale_fill_discrete(guide = NULL)+
  geom_text(data = filter(tots, Type2 != "Wild/unmarked", Type2 != "Unknown"), aes(x = Type2, label = N), y=2, inherit.aes = FALSE)+ xlab(NULL)+
    stat_summary(fun.y=mean, geom="point", shape = 23, fill = "white", size =4) +
  theme_bw()

ggsave("plots/hardvsoft_distance.tiff", device = "tiff", width =6, height =4)

totsby = group_by(recaps3, Type2, BroodYear) %>%
  summarize(N = n())


ggplot(filter(recaps3, Type != "Wild/unmarked", Type != "Unknown"), aes(x = Type2, y = Distance/1000))+ 
  geom_boxplot(aes(fill = Type2))+
  ylab("Distance from Release Site (km)")+
  scale_fill_discrete(guide = NULL)+
  geom_label(data = filter(totsby, Type2 != "Wild/unmarked", Type2 != "Unknown"), 
             aes(x = Type2, label = paste("N=", N)), y=2, inherit.aes = FALSE)+ xlab(NULL)+
    stat_summary(fun.y=mean, geom="point", shape = 23, fill = "white", size =4) +
  facet_wrap(~BroodYear)+
  theme_bw()


ggsave("plots/hardvsoft_distance_BY.tiff", device = "tiff", width =6, height =4)


```


```{r}
library(ggbeeswarm)

ggplot(filter(recaps3, Type != "Wild/unmarked"), aes(x = Type2, y = Distance))+ 
  geom_quasirandom(aes(color = as.factor(BroodYear), shape = Site))+
  ylab("Distance from Release Site (m)")+
  facet_wrap(~FFyn)+
  scale_fill_discrete(guide = NULL)+
  theme_bw()

```
```{r}

ggplot(filter(recaps3, Type != "Wild", BroodYear != 2024), aes(x = Type2, y = Distance))+ 
  geom_quasirandom(aes(color = Type, shape = Site))+
  ylab("Distance from Release Site (m)")+
  scale_fill_discrete(guide = NULL)+
  facet_wrap(~BroodYear)+
  
  theme_bw()

```
```{r}

tots2023 = filter(recaps3, Type != "Wild/unmarked", BroodYear == 2023) %>%
  group_by(Type) %>%
  summarize(N = n()) %>%
  droplevels()

test23 = filter(recaps3, Type != "Wild/unmarked", BroodYear == 2023, !is.na(Distance)) %>%
  droplevels()

ggplot(test23, aes(x = Type, y = Distance))+ 
  geom_boxplot(aes(fill = Type))+
  ylab("Distance from Release Site (m)")+
  scale_fill_discrete(guide = NULL)+
  geom_label(data = tots2023, aes(x = Type, label = paste("N=", N)), y=5000, inherit.aes = FALSE)+
  theme_bw()

```



Map with color variation by distance, to make sure I got it right.

```{r}
recapsf = st_as_sf(filter(recaps3, !is.na(`Longitude`)), coords = c("Longitude", "Latitude"), crs = 4326)


ggplot()+
  geom_sf(data = WW_Delta)+
  geom_sf(data = recapsf, aes(color = Distance), size =3)+
  geom_sf(data = releasesf, aes(shape = Type), size = 5)+
  scale_color_viridis_c(option = "B")+
  coord_sf(xlim = c(-121.5, -122.2), ylim = c(37.8, 38.5))

recapsum = group_by(recaps3, Site, Type, SiteType) %>%
  summarize(Nsmelt = n(), MeanDist = mean(Distance, na.rm = T), sdDist = sd(Distance, na.rm =T))

recapsum
```

```{r}

#ANOVA of differences in distance by release group
library(emmeans)
m1 = lm(Distance ~ Type, data = recaps3)
summary(m1)
anova(m1)
emmeans(m1, pairwise ~ Type)
plot(emmeans(m1, pairwise ~ Type), comparisons = T)

#meh

```

Can cobine Trailer and Large-Scale
Keep carboys seperate
ONly one soft release type

Keep the years togetehr or seperate?

Look at the timing of releases versus number of recaptures. Nov-Dec versus Jan-feb
- Is this all just turbidity? Just at release? or long-term? Pre versus post 'first flush'. We have  a regulatory trigger for when first flush begins. 
- Not all fish are exactly the same age, but they try and realease them all at about the same age, so ones released in November were spawned earlier. 
- Still, check days post hatch

Compare CPUE from EDSM and Salvage

Get data on number released at each event

Soft release in Rio vista in 2022? Check release numbers and timing. 


```{r}

#ANOVA of differences in distance by release event
m1.1 = lm(Distance ~ ReleaseEvent, data = recaps3)
summary(m1.1)
anova(m1.1)
emmeans(m1.1, pairwise ~ ReleaseEvent)
plot(emmeans(m1.1, pairwise ~ ReleaseEvent), comparisons = T)

#It's a bit of a mess.

```


Plot of distance to recapture over time by release type

```{r}

ggplot(recaps3, aes(x = as.numeric(DaysSinceRelease), y = Distance, color = Type))+
  geom_point()+ geom_smooth(method = "lm")+ ylab("Distance from Release Site (m)")+
  xlab("Days since release")

```


Plot of distance to recapture over time by release site

```{r}

ggplot(filter(recaps3, Type == "Hard"), aes(x = as.numeric(DaysSinceRelease), y = Distance, color = Site))+
  geom_point()+ geom_smooth(method = "lm")+ ylab("Distance from Release Site (m)")+
  xlab("Days since release")



```

Do we see more recaptures per released fish during different types of releases?

```{r}


recaps4 = group_by(recaps3, ReleaseEvent, ReleaseDate, Site, Type, Type2, BroodYear) %>%
  summarize(N = n())

releasestat = left_join(release, recaps4) %>%
  filter(Type != "Unknown") %>%
  mutate(BroodYear = as.numeric(str_sub(ReleaseEvent, 3L, end = 6L))) %>%
  left_join(ff) %>%
  mutate(recapsN = case_when(is.na(N) ~0,
                             TRUE ~N),
         RecapRate = recapsN/TotalReleased,
         FFyn = case_when(ReleaseDate > FirstFlush ~ "After First Flush",
                          TRUE ~ "Before First Flush"),
         Type2 = case_when(Type == "Hard (trailer)" ~ "Hard (large scale)",
                           TRUE ~ Type))

totsr = group_by(releasestat, Type2, FFyn) %>%
  summarize(N = n())

ggplot(releasestat, aes(x = Type2, y = RecapRate*100, fill = Type2))+
  geom_boxplot()+
  facet_wrap(~FFyn)+
  ylab("Percentage Recaptured")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust =1),
        legend.position = "none")+
  xlab(NULL)+
  geom_label(data = totsr, aes(x = Type2, label = paste("N=",N)), y = 0, inherit.aes = FALSE)

ggsave("plots/firstflush.tiff", device = "tiff", width =8, height =5)

ggplot(releasestat, aes(x = Type, y = RecapRate*100, color = Type))+
  geom_quasirandom()+
  ylab("Percentage Recaptured")

```



NOw we move on to distance bewteen recaptured fish. I'm doing it oen release at a tiem again. 

```{r, eval=FALSE}

distances = list(NA)

releasex = filter(release, ReleaseEvent %in% unique(recaps3$ReleaseEvent))
for(i in 1:nrow(releasex)) {
  points = filter(recaps3, ReleaseEvent == releasex$ReleaseEvent[i])
  distances[[i]] = Waterdist(Water_map = spacetools::Delta, Points = points, Latitude_column = `Latitude`,
                    Longitude_column = `Longitude`, PointID_column = FishID)
  names(distances[[i]]) = releasex$ReleaseEvent[i]
    
}

# RVhdist = Waterdist(Water_map = spacetools::Delta, Points = RVhardt, Latitude_column = `Latitude`,
#                     Longitude_column = `Longitude`, PointID_column = fishID)
# 
# RVhdistx = Waterdist(Water_map = spacetools::Delta, Points = RVhard, Latitude_column = `Latitude`,
#                     Longitude_column = `Longitude`, PointID_column = fishID)
# 
# SCsoft  = Waterdist(Water_map = spacetools::Delta, Points = Dwscsoft, Latitude_column = `Latitude`,
#                        Longitude_column = `Longitude`, PointID_column = fishID)
# 
# SChard = Waterdist(Water_map = spacetools::Delta, Points = Dwschard, Latitude_column = `Latitude`,
#                    Longitude_column = `Longitude`, PointID_column = fishID)

# RVhdist
# RVhdistx
# SCsoft
# SChard

```

Let's look at the mean of the lower triagle of each of these.

```{r}

clustest = lapply(distances, function(x){df = data.frame(Mean = mean(x[lower.tri(x)], na.rm =T), SD = sd(x, na.rm =T))} )%>%
  bind_rows() %>%
  mutate(Release = releasex$ReleaseEvent)


clustest
```

OK, now how to I put these all together to graph them?

```{r}

alldist = lapply(distances, function(x){
  x[lower.tri(x, diag = T)] = NA
  X = as.data.frame(x) %>%
    mutate(Fish1 = row.names(x)) %>%
    pivot_longer(cols = -Fish1, names_to = "Fish2", values_to = "Distance") %>%
  filter(!is.na(Distance))
  return(X)
})

names(alldist) = releasex$ReleaseEvent 

for(i in 1:14) {
  alldist[[i]] = mutate(alldist[[i]], ReleaseEvent = releasex$ReleaseEvent[i])
}

alldist2 = bind_rows(alldist) %>%
  filter(ReleaseEvent != "NA") %>%
  left_join(release)

totsX = group_by(recaps3, ReleaseEvent) %>%
  summarize(N = n())%>%
  filter(ReleaseEvent != "NA")
ggplot(alldist2, aes(x = ReleaseEvent, y = Distance, fill = ReleaseEvent)) + geom_boxplot()+
  ylab("Distance Between Recaptures")+ theme_bw()+ scale_fill_discrete(guide = NULL)+
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust =1))+
  geom_label(data = totsX, aes(x = ReleaseEvent, y =1, label = paste("N=", N)), fill= "white")

```

```{r}
alldist2.1 = filter(alldist2, Type != "Unknown") %>%
  mutate(BroodYear = case_when(ReleaseDate < ymd("2023-09-01") ~ 2022,
                               TRUE ~ 2023),
         Type2 = case_when(Type == "Hard (trailer)" ~ "Hard (large scale)",
                           TRUE ~ Type))

ggplot(alldist2.1, aes(x = Type2, y = Distance, fill = Type2)) + geom_boxplot()+
  ylab("Distance Between Recaptures (m)")+
  xlab("Release Type")+
   stat_summary(fun.y=mean, geom="point", shape = 23, fill = "white", size =4) +
  theme_bw() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust =1),
        legend.position = "none")

ggsave("plots/DistanceBetweenRecaps.tiff", device = "tiff", width =5, height =4)

ggplot(alldist2.1, aes(x = Type2, y = Distance, fill = Type2)) + geom_boxplot()+
  ylab("Distance Between Recaptures (m)")+
  xlab("Release Type")+
  facet_wrap(~BroodYear)+
   stat_summary(fun.y=mean, geom="point", shape = 23, fill = "white", size =4) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust =1),
        legend.position = "none")

ggsave("plots/DistanceBetweenRecaps_byyear.tiff", device = "tiff", width =8, height =5)

```


now let's look at it statistically

```{r}

m2 = lm(Distance ~ Type, data = alldist2)
summary(m2)
plot(m2)
#QQ plot is a bit of a mess I tried log-transforming and sqrt transforming and its still a mess. 

emmeans(m2, pairwise ~ Type)
plot(emmeans(m2, pairwise ~ Type), comparisons = T)

```

So soft were closer together than hard

I could also do this by release rather than release type


The before/after first flush thing was interesting, something with turbidity for the week after the release?


So, first i need to pull turbidity data from that area for the week after each realease.

```{r}
library(cder)
library(zoo)

turbdat = cdec_query("RVB", sensors = 27, start.date = as.Date("2021-01-01"), end.date = today())

turbdat = mutate(turbdat, Month = month(DateTime), Date = date(DateTime), Year = year(DateTime))

turbdat_ave = group_by(turbdat, Date, Month, Year) %>%
  summarize(Turbidity = mean(Value, na.rm =T))

test = rollmean(turbdat_ave$Turbidity, 7, align = "right", na.pad = TRUE)

turbdat_ave$weekturb = test

releasestat = left_join(releasestat, turbdat_ave, by = c("ReleaseDate" = "Date"))
  
ggplot(filter(releasestat, BroodYear != 2021, Type2 != "NA"), aes(x = weekturb, y = RecapRate )) + geom_point(aes(color = as.factor(BroodYear), shape = Type2), size =3)+ geom_smooth(method = "lm")+ylab("Percentage Recaptured") + 
  xlab("Mean turbidity for 7 days post release (NTU)")+
  theme_bw()+
  scale_color_brewer(palette = "Set2", name = "Brood Year")+
  scale_shape(name = "Release Type")

ggsave("plots/turbidityrecaps.tiff", device = "tiff", width =5, height =4)

```

## Pretty map of recaputres for presentations

```{r}


library(tidyverse)
library(lubridate)
library(sf)
library(ggmap)
library(readxl)
library(deltamapr)
library(ggspatial)

###########################################################
#map for Katie
smelt2 = read_excel("data/Running Delta Smelt Catch.xlsx", 
                    sheet = "Delta Smelt Catch Data") 

smeltsf = st_as_sf(smelt2, coords = c("LongitudeStart", "LatitudeStart"), crs = 4326) %>%
  mutate(Date = str_sub(as.character(SampleDate), 6, 10)) %>%
  filter(SampleDate > ymd("2023-09-30"), !is.na(ReleaseMethod))

ggplot()+
  geom_sf(data = WW_Delta, fill = "grey90")+
  geom_sf(data = smeltsf, aes(shape = ReleaseMethod,  fill = ReleaseMethod), size =3)+
  scale_fill_brewer(palette = "Set2", labels = c("Hard (carboy)", "Hard (large scale)", "Hard (trailer)",
                                                  "NA - unmarked fish", "Soft (carboy)"))+
  scale_shape_manual(values = c(21,22,23,24,25), labels = c("Hard (carboy)", "Hard (large scale)", "Hard (trailer)",
                                                  "NA - unmarked fish", "Soft (carboy)"))+
  #geom_sf_label(data = smeltsf, aes(label = Date), hjust =0, nudge_x = 0.01)+
  coord_sf(ylim = c(37.8, 38.35), xlim = c(-122.2, -121.45))+
  theme_bw()+
  ylab(NULL)+
  xlab(NULL)+
  annotation_scale()+
  annotation_north_arrow(aes(location = "tl"))

ggsave("plots/Releasemap.png", device = "png", width =8, height =6)
```

